Noise-Protected Quantum Memory Registers with Nitrogen-Vacancy Centers and $^{13}$C Nuclear Spins

### 2016-06-23 swap-nvst-paper/paper.tex
added text with an intuitive summary of H_int_dec

Dynamic decoupling schemes effectively prepend a modulation function
\begin{align}
  F\p{t}=\sum_kf_k\cos\p{k\omega_{DD}t-\phi_{DD}}
\end{align}
with fundamental frequency $\omega_{DD}$, phase $\phi_{DD}$, and range
$\set{1,-1}$ to $\sigma_z^\NV$, so that the system Hamiltonian reads
\begin{align}
  \tilde H_\t{int} =
  F\p{t}m_sI_\NV^z\sum_j\tilde{\v A_j}\p{t}\c\v I_j.
\end{align}
At a resonance $k_{DD}\omega_{DD}\equiv\omega_r\gg A_j$, the secular
approximation allows one to neglect $\v A_j^\parallel$ and express
\begin{align}
  \tilde H_\t{int} = \f12 f_{k_{DD}} m_sI_\NV^z
  \sum_{j:\omega_j=\omega_r} A_j^\perp\v I_j\c\uv n_j,
  \label{H_int_resonance}
\end{align}
where $\uv n_j$ is the direction of $\v A_j^\perp$ rotated about
$\uv\omega_j$ by $\phi_{DD}$, i.e.
\begin{align}
  \uv n_j = \uv A_j^\perp\cos\phi_{DD} +
  \uv A_j^{\perp\perp}\sin\phi_{DD}.
\end{align}
One can concisely write (\ref{H_int_resonance}) with
$h_j^\t{int}\equiv f_{k_{DD}}m_sA_j^\perp/2$ and the Pauli spin-1/2
vector $\v\sigma^j$ (as we consider only spin-1/2 nuclei) as
\begin{align}
  \tilde H_\t{int} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\uv n_j.
  \label{H_int}
\end{align}

@@

When $\omega_\t{ctl}\gg\gamma_jB_\t{ctl}$, the secular approximation
allows one to neglect $\v B_{\t{ctl},j}^\parallel$ and express
\begin{align}
  \tilde H_\t{ctl} = -\f12\sum_{j:\omega_j=\omega_\t{ctl}}
  \gamma_jB_{\t{ctl},j}^\perp \v I_j\c\uv m_j,
  \label{H_ctl_resonance}
\end{align}
where $\uv m_j$ is the direction of $\v B_{\t{ctl},j}^\perp$ rotated
about $\uv\omega_j$ by $\phi_\t{ctl}$, i.e.
\begin{align}
  \uv m_j = \uv B_{\t{ctl},j}^\perp\cos\phi_\t{ctl} +
  \uv B_{\t{ctl},j}^{\perp\perp}\sin\phi_\t{ctl}.
\end{align}
One can concisely write (\ref{H_ctl_resonance}) with
$h_j^\t{ctl}\equiv-\gamma_jB_{\t{ctl},j}^\perp/2$ as
\begin{align}
  \tilde H_\t{ctl} = \sum_{j:\omega_j=\omega_\t{ctl}}
  h_j^\t{ctl}\v I_j\c\uv m_j.
  \label{H_ctl}
\end{align}

In both (\ref{H_int}) and (\ref{H_ctl}), one can choose $\phi_{DD}$
and $\phi_\t{ctl}$ such that $\uv n_j\p\phi$ and
$\uv m_j\p{\phi_\t{ctl}}$ lie anywhere in the plane perpendicular to
$\uv\omega_j$. One can also tune $h_j^\t{int}$ and $h_j^\t{ctl}$
respectively via free parameters in the AXY-$n$ sequence and
$\v B_\t{ctl}$. These controls, together with rotations of the NV
electron spin, in principle allow one to realize a universal set of
quantum gates. An exception to this statement is when two or more
nuclei have the same effective larmor frequency $\omega_j$, in which
case they are addressed simultaneously in (\ref{H_int}) and
(\ref{H_ctl}).

@@

In the frame of $\tilde H_\t{dec}$, the entire system is governed by
the Hamiltonian
\begin{align}
  \tilde H_\t{int}^\t{dec} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\tilde{\v n}_j\p{t},
  \label{H_int_dec_full}
\end{align}
where $\tilde{\v n}_j\p{t}$ is the unit vector $\uv n_j$ rotated
about $\uv\nu_j$ by $\nu_jt$, i.e.
\begin{align}
  \tilde{\v n}_j\p{t} = \v n_j^\parallel +
  \v n_j^\perp\cos\p{\nu_jt} + \v n_j^{\perp\perp}\sin\p{\nu_jt},
\end{align}
where
\begin{align}
  \v n_j^\parallel &\equiv \p{\uv n_j\c\uv\nu_j}\uv\nu_j, \\
  \v n_j^\perp &\equiv \uv n_j - \uv n_j^\parallel, \\
  \v n_j^{\perp\perp} &\equiv \uv\nu_j\times\uv n_j.
\end{align}
If $\omega_j\gg\nu_j\gg h_j^\t{int}$, the secular approximation allows
us to neglect $\v n_j^\perp$ and $\v n_j^{\perp\perp}$ in
(\ref{H_int_dec_full}), leaving
\begin{align}
  \tilde H_\t{int}^\t{dec}
  = \sum_{j:\omega_j=\omega_r}
  \tilde h_j^\t{int}I_\NV^z\v I_j\c\uv\nu_j,
  \label{H_int_dec}
\end{align}
where
\begin{align}
  \tilde h_j^\t{int} = h_j\uv n_j\c\uv\nu_j.
\end{align}

The utility of the decoupling field lies in the fact that when
$\uv n_j$ and $\uv\nu_j$ are orthogonal, the interaction with spin $j$
is suppressed as $\tilde h_j^\t{int}=0$. Considering a single larmor
pair, one can therefore apply a decoupling field to eliminate one of
the two terms in (\ref{H_int_dec}) and realize an effective
Hamiltonian which couples the NV electron spin to only one nuclear
spin. This scheme fails for larmor pairs with mutually parallel
$\v A_j^\perp$.

>>>>

Dynamic decoupling schemes effectively prepend a modulation function
\begin{align}
  F\p{t}=\sum_kf_k\cos\p{k\omega_{DD}t-\phi_{DD}}
\end{align}
with fundamental frequency $\omega_{DD}$, phase $\phi_{DD}$, and range
$\set{1,-1}$ to $\sigma_z^\NV$, so that the system Hamiltonian reads
\begin{align}
  \tilde H_\t{int} =
  F\p{t}m_sI_\NV^z\sum_j\tilde{\v A_j}\p{t}\c\v I_j.
\end{align}
At a resonance $k_{DD}\omega_{DD}\equiv\omega_r\gg A_j$, the secular
approximation allows one to neglect $\v A_j^\parallel$ and express
\begin{align}
  \tilde H_\t{int} = \f12 f_{k_{DD}} m_sI_\NV^z
  \sum_{j:\omega_j=\omega_r} A_j^\perp\v I_j\c\uv n_j,
  \label{H_int_resonance}
\end{align}
where $\uv n_j$ is a unit vector in the direction of $\v A_j^\perp$
rotated about $\uv\omega_j$ by $\phi_{DD}$, i.e.
\begin{align}
  \uv n_j = \uv A_j^\perp\cos\phi_{DD} +
  \uv A_j^{\perp\perp}\sin\phi_{DD}.
\end{align}
One can concisely write (\ref{H_int_resonance}) with
\begin{align}
h_j^\t{int}\equiv f_{k_{DD}}m_sA_j^\perp/2
\end{align}
as
\begin{align}
  \tilde H_\t{int} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\uv n_j.
  \label{H_int}
\end{align}

@@

When $\omega_\t{ctl}\gg\gamma_jB_\t{ctl}$, the secular approximation
allows one to neglect $\v B_{\t{ctl},j}^\parallel$ and express
\begin{align}
  \tilde H_\t{ctl} = -\f12\sum_{j:\omega_j=\omega_\t{ctl}}
  \gamma_jB_{\t{ctl},j}^\perp \v I_j\c\uv m_j,
  \label{H_ctl_resonance}
\end{align}
where $\uv m_j$ is the direction of $\v B_{\t{ctl},j}^\perp$ rotated
about $\uv\omega_j$ by $\phi_\t{ctl}$, i.e.
\begin{align}
  \uv m_j = \uv B_{\t{ctl},j}^\perp\cos\phi_\t{ctl} +
  \uv B_{\t{ctl},j}^{\perp\perp}\sin\phi_\t{ctl}.
\end{align}
One can concisely write (\ref{H_ctl_resonance}) with
\begin{align}
h_j^\t{ctl}\equiv-\gamma_jB_{\t{ctl},j}^\perp/2
\end{align}
as
\begin{align}
  \tilde H_\t{ctl} = \sum_{j:\omega_j=\omega_\t{ctl}}
  h_j^\t{ctl}\v I_j\c\uv m_j.
  \label{H_ctl}
\end{align}

In both (\ref{H_int}) and (\ref{H_ctl}), one can choose $\phi_{DD}$
and $\phi_\t{ctl}$ such that $\uv n_j\p\phi$ and
$\uv m_j\p{\phi_\t{ctl}}$ lie anywhere in the plane perpendicular to
$\uv\omega_j$. One can also tune $h_j^\t{int}$ and $h_j^\t{ctl}$
respectively via free parameters in the AXY-$n$ sequence and
$\v B_\t{ctl}$. These controls, together with rotations of the NV
electron spin, allow one to realize a universal set of quantum
gates. An exception to this statement is when two or more nuclei have
the same effective larmor frequency $\omega_j$, in which case they are
addressed simultaneously in (\ref{H_int}) and (\ref{H_ctl}).

@@

In the frame of $\tilde H_\t{dec}$, the entire system is governed by
the Hamiltonian
\begin{align}
  \tilde H_\t{int}^\t{dec} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\tilde{\v n}_j\p{t},
  \label{H_int_dec_full}
\end{align}
where $\tilde{\v n}_j\p{t}$ is the unit vector $\uv n_j$ rotated
about $\uv\nu_j$ by $\nu_jt$, i.e.
\begin{align}
  \tilde{\v n}_j\p{t} = \v n_j^\parallel +
  \v n_j^\perp\cos\p{\nu_jt} + \v n_j^{\perp\perp}\sin\p{\nu_jt},
\end{align}
with
\begin{align}
  \v n_j^\parallel &\equiv \p{\uv n_j\c\uv\nu_j}\uv\nu_j, \\
  \v n_j^\perp &\equiv \uv n_j - \uv n_j^\parallel, \\
  \v n_j^{\perp\perp} &\equiv \uv\nu_j\times\uv n_j.
\end{align}
If $\omega_j\gg\nu_j\gg h_j^\t{int}$, the secular approximation allows
us to neglect $\v n_j^\perp$ and $\v n_j^{\perp\perp}$ in
(\ref{H_int_dec_full}), leaving
\begin{align}
  \tilde H_\t{int}^\t{dec}
  = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\v n_j^\parallel
  = \sum_{j:\omega_j=\omega_r}
  \tilde h_j^\t{int}I_\NV^z\v I_j\c\uv\nu_j,
  \label{H_int_dec}
\end{align}
where
\begin{align}
  \tilde h_j^\t{int} = h_j\uv n_j\c\uv\nu_j.
\end{align}

In words, the decoupling field $\v B_\t{dec}^{\p\omega}\p{t}$ modifies
$\tilde H_\t{int}$ in (\ref{H_int}) by suppressing the components of
$\uv n_j$ (i.e. $\uv A_j^\perp$ rotated about $\uv\omega_j$ by
$\phi_{DD}$) which are perpendicular to the respective $\uv\nu_j$
(i.e.  $\uv B_{\t{dec},j}^\perp$ rotated about $\uv\omega_j$ by
$\phi_\t{dec}$). The utility of this decoupling field lies in the fact
that when $\uv n_j$ and $\uv\nu_j$ are orthogonal, the interaction
with spin $j$ is suppressed as $\tilde h_j^\t{int}=0$. Considering a
single larmor pair, one can therefore apply a decoupling field to
eliminate one of the two terms in (\ref{H_int_dec}) and realize an
effective Hamiltonian which couples the NV electron spin to only one
nuclear spin. This scheme fails for larmor pairs with mutually
parallel $\v A_j^\perp$.
